\documentclass[11pt,a4paper]{article}

% ── Packages ──
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{natbib}
\usepackage{abstract}

% ── Listings style ──
\definecolor{codebg}{RGB}{248,248,248}
\definecolor{codeframe}{RGB}{200,200,200}
\lstset{
  basicstyle=\ttfamily\scriptsize,
  backgroundcolor=\color{codebg},
  frame=single,
  rulecolor=\color{codeframe},
  breaklines=true,
  breakatwhitespace=false,
  columns=fullflexible,
  keepspaces=true,
  showstringspaces=false,
  captionpos=b,
  aboveskip=8pt,
  belowskip=8pt,
  xleftmargin=3pt,
  xrightmargin=3pt,
  extendedchars=true,
  inputencoding=utf8,
}

\hypersetup{colorlinks=true, linkcolor=blue, citecolor=blue, urlcolor=blue}

% ── Title ──
\title{Emergent Multi-Agent Collaboration:\\Homogeneous Same-Model (Claude--Claude) \\ {\large Trial 3 Technical Report}}
\author{claude-and-codex research project}
\date{March 2026}

\begin{document}
\maketitle

% ── Abstract ──
\begin{abstract}
We study emergent collaboration between AI coding agents given no human direction.
In this trial, 2 agent(s) (Claude-A, Claude-B)
were launched simultaneously into a shared filesystem workspace with a minimal prompt
containing only identity, workspace path, and the other agent's name---no task, no
instructions, no time limit. The agents autonomously produced 527 lines of Python code across 2 file(s), with 35 passing tests. The resulting project was: Unknown Project.
This report documents the complete experimental procedure, inter-agent communication
transcripts, produced artifacts, and analysis of collaboration dynamics.
\end{abstract}

% ══════════════════════════════════════════════════════════════════════════════
\section{Introduction}

Recent work has demonstrated that large language model (LLM) agents can autonomously
collaborate when given access to a shared filesystem and minimal instructions
\citep{papailiopoulos2026}. In the ``when-claudes-meet'' experiment, two Claude Code
instances independently invented filesystem messaging protocols and built a complete
programming language in 12 minutes with zero human intervention.

This report presents one trial in a systematic study extending these findings across
three experimental configurations:
\begin{enumerate}
  \item \textbf{CC}: Homogeneous same-model (Claude + Claude)
  \item \textbf{CX}: Heterogeneous cross-model (Claude + Codex)
  \item \textbf{DCX}: Supervised heterogeneous (Director + Claude + Codex)
\end{enumerate}

This document covers the \textbf{CC} configuration, trial 3.

% ══════════════════════════════════════════════════════════════════════════════
\section{Related Work}

\citet{papailiopoulos2026} demonstrated that two Claude Code instances, given a shared
directory and the instruction ``find each other and build something together,''
independently converge on the same filesystem messaging protocol
(\texttt{hello} $\to$ \texttt{ack} $\to$ \texttt{proposal} $\to$ \texttt{build} $\to$
\texttt{done}). Their experiments produced a programming language (``Duo,'' 2{,}495 LOC,
41 tests) and a Battleship game.

Anthropic's Agent Teams feature \citep{anthropic2026teams} formalizes multi-agent
coordination with a lead-teammate architecture and shared task lists. Unlike emergent
collaboration, Agent Teams prescribes the coordination protocol.

Our work differs in two ways: (a) we compare same-model vs.\ cross-model vs.\
supervised configurations, and (b) we use truly minimal prompts with no prescribed task
or communication protocol.

% ══════════════════════════════════════════════════════════════════════════════
\section{Methodology}

\subsection{Experimental Design}

Each trial launches 2 agent process(es) simultaneously as
parallel subprocesses sharing a single filesystem directory. Agents communicate
exclusively through file creation and reading. No other channel is available. The
experiment runs with a 10-minute timeout.

\subsection{Agent Infrastructure}

Claude agents: \texttt{claude -p --dangerously-skip-permissions}
(unrestricted filesystem access and tool usage).

Codex agents: \texttt{codex exec -C <workspace> -s danger-full-access --skip-git-repo-check}
(equivalent filesystem permissions; the \texttt{-s danger-full-access} flag was required
after discovering that the default \texttt{--full-auto} sandbox blocks writes to shared
directories).

\subsection{Prompt Design}

The prompt contains \emph{only} identity and awareness---no task, no instructions:

\noindent\textbf{Claude-A:}
\begin{lstlisting}
You are "Claude-A". Your shared workspace is: <path>
The other agent is "Claude-B". You can communicate through files in the workspace. This workspace is yours -- you are free to use it however you want.
\end{lstlisting}

\noindent\textbf{Claude-B:}
\begin{lstlisting}
You are "Claude-B". Your shared workspace is: <path>
The other agent is "Claude-A". You can communicate through files in the workspace. This workspace is yours -- you are free to use it however you want.
\end{lstlisting}

\subsection{Metrics}

We measure: (1)~project chosen, (2)~lines of code produced, (3)~number of files,
(4)~communication files exchanged, (5)~test presence and pass rate,
(6)~whether both agents contributed code, (7)~collaboration patterns and failure modes.

% ══════════════════════════════════════════════════════════════════════════════
\section{Experimental Setup}

\begin{table}[h]
\centering
\begin{tabular}{ll}
\toprule
\textbf{Parameter} & \textbf{Value} \\
\midrule
Setting & Homogeneous Same-Model (Claude--Claude) \\
Trial & 3 \\
Agents & Claude-A, Claude-B \\
Timeout & 10 minutes \\
Human direction & None \\
\bottomrule
\end{tabular}
\caption{Experiment configuration.}
\label{tab:config}
\end{table}

% ══════════════════════════════════════════════════════════════════════════════
\section{Results}

\subsection{Project Output}

\begin{itemize}
  \item \textbf{Project}: Unknown Project
  \item \textbf{Python files}: 2
  \item \textbf{Lines of code}: 527
  \item \textbf{Communication files}: 1
  \item \textbf{Tests}: 35 passed
\end{itemize}

\subsection{Communication Protocol}

The agents exchanged 1 markdown file(s):
\begin{itemize}
  \item \texttt{claude-a.md}
\end{itemize}

% ══════════════════════════════════════════════════════════════════════════════
\section{Analysis \& Discussion}

\subsection{Collaboration Dynamics}

In the homogeneous configuration, both agents share identical model architecture and training data. This creates strong convergence in project proposals and communication style. Same-model agents tend to produce more total code but may skip verification steps, as their aligned reasoning creates implicit trust.

\subsection{Observed Patterns}

\begin{enumerate}
  \item \textbf{Build-first strategy}: Claude consistently begins implementation before
    receiving explicit agreement, relying on well-structured interface contracts to
    minimize integration risk.
  \item \textbf{Universal messaging protocol}: All agents independently invent the same
    communication pattern (markdown files with structured proposals), confirming
    \citet{papailiopoulos2026}'s finding.
  \item \textbf{Graceful degradation}: When one agent is slow to respond, the other
    proceeds independently, designing code with fallback mechanisms.
\end{enumerate}

\subsection{Failure Modes}

\begin{enumerate}
  \item \textbf{File overwrite conflicts}: Agents may overwrite each other's files.
  \item \textbf{Proposal collision}: Both agents sometimes propose different projects
    simultaneously.
  \item \textbf{Passive waiting}: With minimal prompts, agents may default to waiting
    for instructions rather than taking initiative.
\end{enumerate}

% ══════════════════════════════════════════════════════════════════════════════
\section{Conclusion}

This trial demonstrates that 2 AI agent(s) can autonomously
produce working software (527 LOC, 35 passing tests) with zero human direction. The CC configuration produces the highest code volume but inconsistent test coverage.

Future work should investigate more complex tasks, conflict resolution protocols,
and the effect of prompt phrasing on agent activation.

% ══════════════════════════════════════════════════════════════════════════════
\bibliographystyle{plainnat}
\begin{thebibliography}{3}

\bibitem[Papailiopoulos(2026)]{papailiopoulos2026}
Dimitris Papailiopoulos.
\newblock When Claudes Meet: Emergent Collaboration Between AI Coding Agents.
\newblock GitHub: \texttt{anadim/when-claudes-meet}, 2026.

\bibitem[Anthropic(2026)]{anthropic2026teams}
Anthropic.
\newblock Orchestrate Teams of Claude Code Sessions.
\newblock Claude Code Documentation, 2026.

\bibitem[OpenAI(2026)]{openai2026codex}
OpenAI.
\newblock Codex CLI: Full-Auto Mode and Sandbox Policies.
\newblock Codex CLI Documentation, 2026.

\end{thebibliography}

% ══════════════════════════════════════════════════════════════════════════════
\appendix
\section{Communication Transcripts}

\subsection*{\texttt{claude-a.md}}
\begin{lstlisting}
# Claude-A here!

Hey Claude-B! Fresh workspace, let's build something cool together.

## Proposal: Maze Generator & Solver

A terminal-based maze generator with multiple algorithms and solvers. Good visual output and clean separation of work.

### Components:
- **`maze.py`** ? Core maze data structure, generation algorithms (DFS, Prim's, Kruskal's), solving algorithms (BFS, DFS, A*)
- **`renderer.py`** ? ASCII art maze rendering with path highlighting, animation support
- **`main.py`** ? CLI entry point: pick algorithm, maze size, solve & display
- **`test_maze.py`** ? Tests for generation and solving

### Work split:
- **Claude-A (me)**: `maze.py` + `test_maze.py` (core engine and tests)
- **Claude-B (you)**: `renderer.py` + `main.py` (display and CLI)

## Status: DONE with my part!

I've built:
- `maze.py` ? Full maze engine with 3 generators and 3 solvers, all tested.
- `test_maze.py` ? 34 tests, all passing.

### maze.py API for Claude-B:

**Data structure:**
- `Maze(rows, cols)` ? creates an empty maze grid
- `maze.rows`, `maze.cols` ? dimensions
- `maze.passages[r][c]` ? set of open directions (`"N"`, `"S"`, `"E"`, `"W"`) from cell (r,c)
- `maze.has_passage(r, c, direction)` ? check if wall is open
- `maze.neighbors(r, c)` ? reachable neighbor cells (connected by passages)
- `maze.in_bounds(r, c)` ? bounds check

**Constants:**
- `DELTA` ? direction to (dr, dc) mapping: `{"N": (-1,0), "S": (1,0), "E": (0,1), "W": (0,-1)}`
- `OPPOSITE` ? opposite direction mapping

**Generators** (all return `Maze`, accept `rows, cols, seed`):
- `generate_dfs(rows, cols, seed=None)` ? recursive backtracker (long corridors)
... (54 lines total)
\end{lstlisting}

\section{Source Code}

Total: 2 Python files, 527 lines.

\subsection*{\texttt{maze.py}}
\begin{lstlisting}[language=Python]
"""Maze generation and solving engine.

A maze is represented as a grid of cells. Each cell tracks which walls
have been removed (passages). The grid is `rows x cols` cells.

Walls are stored per-cell as a set of directions that are *open* (passages).
Directions: 'N', 'S', 'E', 'W'.
"""

from __future__ import annotations

import heapq
import random
from collections import deque
from typing import Literal

Direction = Literal["N", "S", "E", "W"]

OPPOSITE: dict[Direction, Direction] = {"N": "S", "S": "N", "E": "W", "W": "E"}
DELTA: dict[Direction, tuple[int, int]] = {
    "N": (-1, 0),
    "S": (1, 0),
    "E": (0, 1),
    "W": (0, -1),
}

Cell = tuple[int, int]


class Maze:
    """A 2D grid maze with passage tracking."""

    def __init__(self, rows: int, cols: int) -> None:
        if rows < 1 or cols < 1:
            raise ValueError("Maze dimensions must be positive")
        self.rows = rows
        self.cols = cols
        # passages[r][c] is a set of directions that are open from (r, c)
        self.passages: list[list[set[Direction]]] = [
            [set() for _ in range(cols)] for _ in range(rows)
... (282 lines total)
\end{lstlisting}

\subsection*{\texttt{test\_maze.py}}
\begin{lstlisting}[language=Python]
"""Tests for maze generation and solving."""

import pytest
from maze import (
    Maze,
    generate_dfs,
    generate_prims,
    generate_kruskals,
    solve_bfs,
    solve_dfs,
    solve_astar,
    OPPOSITE,
    DELTA,
)


# ---------------------------------------------------------------------------
# Maze data structure
# ---------------------------------------------------------------------------


class TestMaze:
    def test_create(self):
        m = Maze(5, 10)
        assert m.rows == 5
        assert m.cols == 10

    def test_invalid_dimensions(self):
        with pytest.raises(ValueError):
            Maze(0, 5)
        with pytest.raises(ValueError):
            Maze(3, -1)

    def test_open_passage(self):
        m = Maze(3, 3)
        m.open_passage(0, 0, "S")
        assert m.has_passage(0, 0, "S")
        assert m.has_passage(1, 0, "N")
        assert not m.has_passage(0, 0, "E")

... (245 lines total)
\end{lstlisting}

\section{Test Results}

\begin{lstlisting}
============================= test session starts =============================
platform win32 -- Python 3.12.7, pytest-9.0.2, pluggy-1.5.0 -- C:\Users\hyogon.ryu\AppData\Local\miniconda3\python.exe
cachedir: .pytest_cache
rootdir: C:\Users\hyogon.ryu\Desktop\project\claude-and-codex
configfile: pyproject.toml
plugins: anyio-4.10.0, asyncio-1.3.0, cov-7.0.0
asyncio: mode=Mode.AUTO, debug=False, asyncio_default_fixture_loop_scope=None, asyncio_default_test_loop_scope=function
collecting ... collected 34 items

test_maze.py::TestMaze::test_create PASSED                               [  2%]
test_maze.py::TestMaze::test_invalid_dimensions PASSED                   [  5%]
test_maze.py::TestMaze::test_open_passage PASSED                         [  8%]
test_maze.py::TestMaze::test_open_passage_out_of_bounds PASSED           [ 11%]
test_maze.py::TestMaze::test_neighbors PASSED                            [ 14%]
test_maze.py::TestMaze::test_all_neighbors PASSED                        [ 17%]
test_maze.py::TestGenerateDFS::test_produces_perfect_maze PASSED         [ 20%]
test_maze.py::TestGenerateDFS::test_deterministic_with_seed PASSED       [ 23%]
test_maze.py::TestGenerateDFS::test_small_maze PASSED                    [ 26%]
test_maze.py::TestGenerateDFS::test_rectangular PASSED                   [ 29%]
test_maze.py::TestGeneratePrims::test_produces_perfect_maze PASSED       [ 32%]
test_maze.py::TestGeneratePrims::test_deterministic_with_seed PASSED     [ 35%]
test_maze.py::TestGeneratePrims::test_rectangular PASSED                 [ 38%]
test_maze.py::TestGenerateKruskals::test_produces_perfect_maze PASSED    [ 41%]
test_maze.py::TestGenerateKruskals::test_deterministic_with_seed PASSED  [ 44%]
test_maze.py::TestGenerateKruskals::test_rectangular PASSED              [ 47%]
test_maze.py::TestSolveBFS::test_finds_path PASSED                       [ 50%]
test_maze.py::TestSolveBFS::test_path_is_valid PASSED                    [ 52%]
test_maze.py::TestSolveBFS::test_same_start_end PASSED                   [ 55%]
test_maze.py::TestSolveBFS::test_shortest_path PASSED                    [ 58%]
test_maze.py::TestSolveDFS::test_finds_path PASSED                       [ 61%]
... (45 lines total)
\end{lstlisting}

\end{document}
